---
title: "L03 Tidy Data"
subtitle: "Data Science 1 with R (STAT 301-1)"
author: "YOUR NAME"
pagetitle: "L03 YOUR NAME"
date: today

format:
  html:
    toc: true
    toc-depth: 4
    toc-location: left
    embed-resources: true
    code-fold: false
    link-external-newwindow: true

execute:
  warning: false

from: markdown+emoji
reference-location: margin
citation-location: margin 
---

::: {.callout-tip icon=false}

## Github Repo Link

To link to your github **repo**sitory, appropriately edit the example link below. Meaning replace `https://your-github-repo-url` with your github repo url. Suggest verifying the link works before submitting.

[https://your-github-repo-url](https://your-github-repo-url)

:::

## Overview

The goal of this lab is to learn what it means to be a "tidy" dataset and how to tidy messy datasets utilizing the `tidyr` package -- a core member of the `tidyverse`. 

Useful resources:

- [`tidyr` package home page](http://tidyr.tidyverse.org/)
- [pivoting vignette](https://tidyr.tidyverse.org/articles/pivot.html)

## Load packages

You should always begin by loading all necessary packages towards the beginning of your document.

```{r}
#| label: load-pkgs
#| code-fold: false

# Loading package(s)

```

## Datasets 

All datasets for the Exercises are either defined inline, provided within the core `tidyverse` packages (`table1`, `table2`, `table4a`, `table4b`), or provided in the `fivethirtyeight` package (`drinks`). 

The Case Study dataset is stored in the `data` folder and called `users_top7_2020.csv`. The code to read in the case study data is provided below. We will learn more about reading in data in Chapter 7.

```{r}
#| label: load-data

# Load dataset(s)
users_top7_2020 <- read_csv("data/users_top7_2020.csv")
```

## Exercises

### Exercise 1 

Recreate the plot below showing the change in cases over time using `table2`. What do you need to do first?

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

### Exercise 2

Tidy the simple tibble of [M&M](https://en.wikipedia.org/wiki/M%26M%27s) data below and drop the NA values.

Do you need to make it wider or longer? What are the variables in your tidy version? What argument drops the NA values?

```{r}
#| label: ex-02
#| eval: false

# simple tibble of M&M data
mm_data <- tribble(
  ~mm_type, ~blue, ~orange,	~green,	~yellow, ~brown, ~red, ~cyan_blue,
  "plain",  6,     18,      12,	    6,       7,	     7,    NA,
  "peanut", NA,	   11,	    9,	    1,	     12,	   8,    15
)
```

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

### Exercise 3

Use `table4a` and **only** the `pivot_longer` function to recreate the following:

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

### Exercise 4

Use the `drinks` dataset and **only** the `pivot_longer` function to recreate the following:

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

### Exercise 5

What happens if you use `pivot_wider()` on this table so that we have a dataset with 3 columns (`respondent_name`, `age`, `height`) and why? 

Fix the issue by adding a new column.

```{r}
#| label: ex-05
#| eval: false

# dataset/table
people <- tribble(
  ~respondent_name,  ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

## Case Study

::: {.callout-note collapse=true}

The case study portion for labs is different than the exercises portion. The core learning objective for the case study portion is to see if you can bring tools together and produce work meant for more general consumption. That is, can you bring everything together and answer a question clearly and concisely with supporting evidence.

:::

Tinder is interested in knowing more about their most active users and have tasked you with exploring their 7 most active users during 2020. The dataset containing the top 7 active users during 2020 is stored in the `data` folder and called `users_top7_2020.csv`.^[This dataset was sourced from [Swipestats.io](https://www.swipestats.io/).] Code to read in the data was provided above.

The column names contain prefixes `"matches_"`, `"likes_"`, and `"passes_"` followed by a number; the suffix number represents the month; and the cell represents either the total number of matches (matches), total number of times the user swiped right (likes), or total number of times the user swiped left (passes) during that month (ie: `matches_1` = total number of matches during January 2020).

Use an appropriate graph to visualize the matches, likes, and passes over time for each user. What insights and conclusions can you gain from this graph, if any? Your boss is also taking this opportunity to perform a performance review. Meaning, your boss will also be having your code reviewed for quality. They have instructed you to include the code, but have it folded/hidden from view. This way it won't distract from the final product, but will be accessible for the code reviewers evaluation. 

::: {.callout-warning}

Remember, the final product is meant for your boss and they expect your work to be clear, to the point, and presentable! While they generally don't care about code and wouldn't want it included, they have instructed you to include this time because they are conducting a performance review which includes someone reviewing the quality of your code. Even though you are including the code, they want it folded/hidden from view so not to distract from the final product (use `code-fold: true` in the code block/chunk YAML).

:::

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

## Graduate Challenge

::: callout-important

Graduate students are required to complete this challenge. **It is optional for undergraduate students.**

:::

Use `pivot_wider` to undo the data tidying you applied in the case study exercise above. That is, take the tidied data and turn it back into the original untidy format we had when we read in the `users_top7_2020` dataset.

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::
